---
title: "R Notebook"
output: html_notebook
---

### Class 1: Exploring the ArcGIS Software Suite
In-class Exercise

- Tools
- Set selectable layers
- Attribute manipulation

- **Deliverable**: exported table

Assignment

- Topology exercise


### Class 2: Displaying Data

In-class Exercise

- Data prep
- Exploring existing ArcMap document
- Exploring ArcMap
- Symbology exercise
- Properties exercise
- Graphs in ArcMap

- **Deliverables**: 
  - Exported Symbols .jpeg
  - Exported Propertise .jpeg
  - Exported Graph as .jpeg
  
Assignment

- Symbology and classigication
- ArcMap practice

- **Deliverable**: Exported map with symbolized data

### Class 3: Information Assessment

- Attribute Query

```{r eval=FALSE, include=FALSE, results='hide'}
library(sf)
library(tidyverse)
library(dplyr)
library(raster)
```
  
  
  
```{r}
VA_counties <- st_read(dsn = "Class3_data/VA_Counties.shp")

methods(class = "sf")

head(VA_counties)

##########################################################

plot(VA_counties["COUNTY_NAM"], reset = FALSE, col = "white")

plot(VA_counties %>%
  filter( COUNTY_NAM == "Chesterfield"), 
  col = "red",
  add = TRUE,
  reset = FALSE)

plot(VA_counties %>%
  filter( COUNTY_NAM == "Henrico"), 
  col = "blue",
  add = TRUE)


```
  
- Spatial Query

How many points fall within Chesterfield County?

```{r}
samples <- st_read(dsn = "Class3_data/strmsamplesite_dd83.shp")

plot(VA_counties["COUNTY_NAM"], reset = FALSE, col = "white")
plot(samples["VSS"], add = TRUE)

# data does not overlay --> check projection (should have done this before anything else)

st_crs(samples)
st_crs(VA_counties)
```

  
```{r}

# transforming sample site projection to match the county boundaries

samples_proj <- st_transform(samples, "+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs")
counties_proj <- st_transform(VA_counties, "+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs")

# plot to see if they overlay

plot(counties_proj["COUNTY_NAM"], reset = FALSE, col = "white")
plot(samples_proj["VSS"], add = TRUE)

# Join the county name column from the VA counties data set to the sample site data set

join <- st_join(samples_proj, counties_proj["COUNTY_NAM"])
```
  
```{r}
# number of sample sites in Chesterfield --> filter data, count the number of rows

join %>%
  filter(COUNTY_NAM == "Chesterfield") %>%
  nrow()

# 34 sample sites in Chesterfield
```

How many roads travel through Chesterfield County?

```{r}

roads <- st_read(dsn = "Class1_data/vector/VDOT_Main_Roads.shp")

st_crs(roads)

roads_proj <- st_transform(roads, "+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs")

st_crs(roads_proj)

head(roads)

plot(roads_proj["ROUTE_NAME"])

roads_join <- st_join(counties_proj, roads_proj["ROUTE_NAME"])

roads_join %>%
   filter(COUNTY_NAM == "Chesterfield") %>%
  nrow()

```

  
```{r}
#Join the other way

roads_join2 <- st_join(roads_proj, counties_proj["COUNTY_NAM"])

roads_join2 %>%
   filter(COUNTY_NAM == "Chesterfield") %>%
  nrow()

head(roads_join2)
```
  
```{r}

roads_Ch2 <- roads_join2 %>%
   filter(COUNTY_NAM == "Chesterfield")

plot(counties_proj["COUNTY_NAM"] %>%
  filter( COUNTY_NAM == "Chesterfield"),
  col = "aliceblue",
  reset = FALSE) +
plot(roads_Ch2["ROUTE_NAME"],
     col = "darkslategray",
  add = TRUE)

```

```{r}
# Get rid of roads that extend past county boundary

roads_join3 <- st_join(roads_proj, counties_proj["COUNTY_NAM"], join = st_covered_by)
head(roads_join3)

roads_Ch3 <- roads_join3 %>%
   filter(COUNTY_NAM == "Chesterfield")

plot(counties_proj["COUNTY_NAM"] %>%
  filter( COUNTY_NAM == "Chesterfield"),
  col = "aliceblue",
  reset = FALSE) +
plot(roads_Ch3["ROUTE_NAME"],
     col = "darkslategray",
  add = TRUE)

```
  
  
```{r}

roads_join3 %>%
   filter(COUNTY_NAM == "Chesterfield") %>%
  nrow()

# 919 roads with "covered by" vs 948 roads with "intersects"
```
  
In-class Exercise

- Attribute Query
```{r}
# Using county data and stream sample sites

head(counties_proj)
head(samples_proj)

plot(counties_proj["COUNTY_NAM"],
     col = "cornsilk",
     border = "cornsilk2",
     reset = FALSE) +
plot(samples_proj["VSS"], 
     add = TRUE)

```
-
  
  - How many sample sites have a VSS score greater than 57?
```{r}
samples_proj %>%
  filter(VSS > 57) %>%
  nrow()

#6 48 sample sites have a VSS score greater than 57
```
-

  - How many stream sample sites have a VSS score in the 70s?
```{r}
samples_proj %>%
  filter(VSS > 69) %>%
  filter(VSS < 80) %>%
  summary()

# Use summary to check your work by looking at min and max

# 183 stream sample sites have a VSS score in the 70s
```

- Location Query

  - How many stream sample sites fall within Hanover County?
```{r}

# Spatial join sample sites to counties

join <- st_join(samples_proj, counties_proj["COUNTY_NAM"])

# Use filter to just see Hanover

join %>%
  filter(COUNTY_NAM == "Hanover") %>%
  nrow()

# There are 66 sample sites in Hanover County
```
- Combining Queries

  - How many stream sample site points fall within Accomack County?
```{r}

join %>%
  filter(COUNTY_NAM == "Accomack") %>%
  nrow()

# There are 35 stream sample sites in Accomack County

```
- 
  
  - Of those stream sample site points that fall within Accomack, how many have a VSS score greater than 69?
  
```{r}

join %>%
  filter(COUNTY_NAM == "Accomack") %>%
  filter(VSS > 69) %>%
  nrow()

# 3 of the stream sample sites in Accomack County have a VSS score greater than 69
```
  
 - Joining tables
 
  - Join the stream sample sites_Join.dbf to the stream sample site points based off of the Loc_Code field
```{r}

library("foreign")

sample_dbf <- read.dbf("Class3_data/strmsamplesites_Join.dbf")

dbf_join <- merge(samples_proj, sample_dbf, by="LOC_CODE")
```

-

  - Symbolize the stream sample site points based on the MIBI scores. Use 5 intervals using natural breaks
  
```{r}
library(RColorBrewer)
plot(counties_proj["COUNTY_NAM"],
     col = "cornsilk1",
     border = "cornsilk4",
     reset = FALSE) +
plot(dbf_join["MIBI_SCORE"],
     breaks = "jenks", nbreaks= 5, 
     pal = brewer.pal(5,"Dark2"),
     pch = 20,
     add = TRUE) 
  

```
  
  #######################
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  