---
title: "Class 5: Cartography"
output: html_notebook
---

### Class 5 Exercise
Make a research publication map to communicate the land cover and stream reaches in your research study area.

Load in Data
```{r echo=TRUE}
# Load in packages I might use
library(sf)
library(tidyverse)
library(dplyr)
library(raster)
library(rgdal)
library(sp)
library(rgeos)
library(RColorBrewer)

# Load in data

NHD_MD <- "Class5_data/NHD_H_Maryland_GDB.gdb" # the input file geodatabase

subset(ogrDrivers(), grepl("GDB", name)) # List all feature classes in a file geodatabse
fc_list <- ogrListLayers(NHD_MD)
print(fc_list) 

NHDFlowline <- readOGR(dsn = NHD_MD, layer = "NHDFlowline") # Flowlines from the National Hydrography Dataset
WBDHU12 <- readOGR(dsn = NHD_MD, layer = "WBDHU12") # Watershed boundaries from the National Hydrography Dataset

Counties <- st_read(dsn = "Class5_data/esri_shps/USCounties_generalized.shp") # Delineates general boundaries of US counties
States <- st_read(dsn = "Class5_data/esri_shps/USStates_generalized.shp") # Delinieates general boundaries of US states
        
```

Project Data if Necessary

```{r echo=TRUE}

projection(NHDFlowline) 
projection(WBDHU12)

# Need to assign projection information (Geographic NAD 1983)
# Google "Geographic NAD 1983 EPSG" --> EPSG:4269

proj4string(NHDFlowline) <- CRS("+init=epsg:4269")
proj4string(WBDHU12) <- CRS("+init=epsg:4269")

# The geodatabase files use "sp" package functions, the .shp files use "sf" package functions

st_crs(Counties)
st_crs(States)

Counties <- st_set_crs(Counties, 4269)
States <- st_set_crs(States, 4269)

```

Study area is the county of Frederick, MD. The map should isolate all data to that boundary

Include the following elements:

- Flowlines from the NHD
- Watershed boundaries from the NHD
- Label the hydrological units
- Include a legend for the NHD flowlines
- Include a scale bar
- Create an inset map and include:
  - Show the state layer
  - Identify the watershed of interest
  - Have a bounding neatline/grid
- Label the figure with a title
- Export a JPEG of the map

```{r echo=TRUE}
# Isolate Frederick County, Maryland

Frederick <- Counties %>%
  filter(STATE_NAME == "Maryland") %>%
  filter(NAME == "Frederick")

plot(Frederick["NAME"])

```

Subset Flowlines and Hydrological Units to Frederick County

```{r echo=TRUE}

NHDFlowline_sf <- st_as_sf(NHDFlowline) # Need to convert flowlines to "sf" object so we can subset
NHDFlowline_sf <- st_set_crs(NHDFlowline_sf, 4269)
Flowline_Frederick <- NHDFlowline_sf[Frederick,]

plot(Flowline_Frederick["FType"])

Flowline_Frederick
```

```{r}

WBDHU12_sf <- st_as_sf(WBDHU12) # Need to convert hydrologic units to "sf" so we can subset
WBDHU12_sf <- st_set_crs(WBDHU12_sf, 4269)
WBDHU12_Frederick <- WBDHU12_sf[Frederick, ] # Subset hydrologic units to Frederick County

```

```{r}
plot(Frederick["NAME"],
     reset=FALSE)
plot(WBDHU12_Frederick["HUC12"],
     add = TRUE)
```



```{r}

plot(Frederick["NAME"],
     col = "blue",
     reset = FALSE)

plot(WBDHU12_Frederick["HUC12"],
     col = "white",
     add = TRUE,
     reset = FALSE)

plot(Flowline_Frederick["FType"],
     add = TRUE)

# I don't think I can clip to the county polygon using subsetting. I think I need to spatial join
```


```{r}
HUC_Fred <- st_intersection(WBDHU12_sf, Frederick)
Flow_Fred <- st_intersection(NHDFlowline_sf, Frederick)

plot(Flow_Fred["FType"])

plot(HUC_Fred["HUC12"],
     col = "white",
     reset = FALSE)
plot(Flow_Fred["FType"],
     col = c("#0a637e","#2591ff","#baa8df","#750000"),
     add = TRUE)

Flow_Fred["FType"]

length(unique(Flow_Fred$FType))
```


```{r}

plot(HUC_Fred["HUC12"],
     border = "black",
     col = "lightyellow",
     reset = FALSE)
plot(Flow_Fred["FType"],
     add = TRUE)
```



```{r}
class(Flow_Fred$FType)
Flow_Fred$FType <- as.factor(Flow_Fred$FType)
nlevels(Flow_Fred$FType)
```

```{r}
plot(States) %>%
  filter(STATE_NAME == "Maryland")

Maryland <- States %>%
  filter(STATE_NAME == "Maryland")

plot(Maryland["STATE_NAME"],
     reset = FALSE)
plot(Frederick["NAME"],
     add = TRUE)
```

```{r}
WBDHU12_Frederick
```







