---
title: Class 6 - Editing
output: html_notebook
---

**Objectives:**

- Create a shapefile
- Edit shapefile data
- Create a geodatabase
- Create a feature class with data

### Create a shapefile exercise

```{r include=FALSE}
# Load in packages I might use
library(sf)
library(tidyverse)
library(dplyr)
library(raster)
library(rgdal)
library(sp)
library(rgeos)
library(RColorBrewer)
library(maps)
library(ggmap)

```

- Create a new line shapefile called roadcenterlines.shp
  - Set projection to Virginia State Plane South NAD 83
  - Add the following fields:
    - Name: this will be the name of the road
    - RdLength: this will be where the road length is calculated
    - CHGBY: this field holds the initials of the analyst that creates this edit. It will be the analysts's initials
    
- Create a new polygon shapefile called Diamond.shp
  - Set projection to Virginia State Plane South NAD 83
  - Add the following fields:
    - Name
    - Acres

```{r echo=TRUE}

# Create a simple dataframe

roads <- data.frame("Name", "RdLength", "CHGBY")

diamond <- data.frame("Name", "Acres")

# These dataframes are empty, we add data in the next step

#-----------------------------------------------------------------

# Locate the Diamond Stadium. The stadium is located at 77° 27’ 48.211 West, 37° 34’ 18.176 North
# In decimal degrees -> -77.46339167, 37.5717167

myMap <- get_stamenmap(bbox = c(left = -77.4675,
                                    bottom = 37.568,
                                    right = -77.46,
                                    top = 37.574),
          maptype = "terrain", 
          crop = TRUE,
          zoom = 16)

ggmap(myMap)










```


```{r echo=TRUE}

library(leaflet)
library(leaflet.extras)


leaflet() %>% setView(lng = -77.46339, lat = 37.57171, zoom = 16) %>% addTiles() %>%
  addCircles(
    lng = -77.46339167, lat = 37.5717167,
    radius = 80) %>%
   addRectangles(
    lng1=-77.4617, lat1=37.5708,
    lng2=-77.463, lat2=37.569) %>%
  addMeasurePathToolbar(options = measurePathOptions(showOnHover = FALSE, minPixelDistance = 30,
showDistances = TRUE, showArea = TRUE, imperial = TRUE)
)

```

```{r}

map <- leaflet() %>% setView(lng = -77.46339, lat = 37.57171, zoom = 16) %>% addTiles() %>%
  enableMeasurePath() %>%
  addDrawToolbar(
    targetGroup = "draw",
    editOptions = editToolbarOptions(selectedPathOptions = selectedPathOptions()
    )
  )  %>%
  addLayersControl(
    overlayGroups = c("draw"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addStyleEditor() %>%
  addMeasurePathToolbar(options = measurePathOptions(showOnHover = FALSE, minPixelDistance = 30,
showDistances = TRUE, showArea = TRUE, imperial = TRUE)
)

map
library(mapview)
library(rgeos)
library(mapedit)
library(lwgeom)

m <- editMap(map)
m
mapview(m$finished)
mf <- m$finished
mf

st_area(mf)

mf$area <- st_area(mf)
mf
mfdf <- as.data.frame(mf)
mfdf

write_rds(mf, path = "diamond.rds")

diamond <- read_rds("diamond.rds")
diamond
```


```{r}
map <- leaflet() %>% 
  setView(lng = -77.46339, lat = 37.57171, zoom = 16) %>%
  addTiles() %>%
  addPolygons(data = diamond)
map

map <- editMap(map)
mapview(map$finished)
map.fin <- map$finished

map.fin$length <- st_length(map.fin)
map.fin$length.ft <- map.fin$length * 3.28084
map.fin


write_rds(map.fin, path = "roads.rds")
roads <- read_rds("roads.rds")
roads$name <- c("Road 1", "Road 2")

map <- leaflet() %>% 
  setView(lng = -77.46339, lat = 37.57171, zoom = 16) %>%
  addTiles() %>%
  addPolygons(data = diamond) %>%
  addPolylines(data = roads, popup = ~as.character(roads$name))
map

mapshot(map, url = "map.html")
```

```{r}

cad <- map <- leaflet() %>% 
  setView(lng = -76.855690, lat = 37.924421, zoom = 18) %>%
  addProviderTiles("Esri.WorldImagery")

cad

library(leafpm)

cad <- editMap(cad, editor = "leafpm")
cad.fin <- cad$finished
write_rds(cad.fin, path = "parcels.rds")
parcels <- read_rds("parcels.rds")


```


```{r}
map <- leaflet() %>% 
  setView(lng = -77.46339, lat = 37.57171, zoom = 10) %>%
  addTiles() %>%
  addPolygons(data = diamond) %>%
  addPolygons(data = parcels) %>%
  addPolylines(data = roads, popup = ~as.character(roads$name))
map
map <- editMap(map, editor = "leafpm")

```


```{r}
parcels <- editFeatures(parcels, editor = "leafpm")
write_rds(parcels, path = "parcels.rds")

read_rds("parcels.rds")
mapview(parcels)
```













